---
title: "FMD modelling"
author: "F.Omata"
date: "2017/10/06"
output:
  pdf_document: 
    latex_engine: lualatex 
documentclass: ltjsarticle 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Plots for FMD Modelling

- Analysis and plot for FMD
- Please run this file with the needed libraries installed and a script "multiplot.R"
- To install package run: `install.packages("package_name")`
- This script will make a directory called "dir" to save all plots

```{r, warning=FALSE, message=FALSE}
rm(list=ls())
dir.create(file.path("./figs"), showWarnings = FALSE)

require(deSolve)
require(ggplot2)
require(dplyr)
require(tidyverse)

source("./multiplot.R")
```

## Initial parameters
latent period cattle: 3.59  
subclinical period cattle: 2.04  
clinical period cattle: 2.35  

latent period swine: 3.07  
subclinical period swine: 2.27  
clinical period swine: 3.42  

R_0: 38.4 *Need verification for parameter*  
cattle to cattle: cattle to swine: swine to cattle: swine to swine = 1:0.77:2.45:3.01

from: http://www.maff.go.jp/j/wpaper/w_maff/h22/pdf/z_1_1_5.pdf  
平成22年度 食料・農業・農村白書（平成23年5月31日公表）  
発生農場は 292 農場、発生自治体数は 11 市町、家畜への被害は牛 69,454 頭、
豚 227,949 頭、その他（山羊、羊、イノシシ、水牛等）405 頭

```{r}
# For cattle
latent_period_cow <- 3.59
subclinical_period_cow <- 2.04
clinical_period_cow <- 2.35 * 20  # most interersting parameter

# For swine
latent_period_swine <- 3.07
subclinical_period_swine <- 2.27
clinical_period_swine <- 3.42 * 20  # most interersting parameter

#depopulation_capacity <- 100
#depopulation_ratio_swine <- 0.4
#vaccination_efficiency <- 0.87

R_0 <- 38.4
beta <- 4.5/(3*2800*5*3.7)  #2.895753e-05

#A <- matrix(0,2,2)
#A[1,] <- c(1,2.45)
#A[2,] <- c(0.77,3.01)
#eigen(A)
params_func <- function(beta.=beta){
    params <- list(
    beta_cowtocow = beta.,
    beta_cowtoswine = 0.77*beta.,
    beta_swinetocow = 2.45*beta.,
    beta_swinetoswine = 3.01*beta.,
    lambda_cow = 1/latent_period_cow,
    lambda_swine = 1/latent_period_swine,
    gamma_cow = 1/subclinical_period_cow,
    gamma_swine = 1/subclinical_period_swine,
    #delta_cow = (1-depopulation_ratio_swine)*depopulation_capacity,
    #delta_swine = depopulation_ratio_swine*depopulation_capacity,
    recover_cow = 1/clinical_period_cow,
    recover_swine = 1/clinical_period_swine,
    day_depopulation_start = 45)  ## need confirmation
    return(params)
    }
```

### Initial conditions  
- From Itao san's email from 2017/10/03  
estimated onset: 04/17  
depopulation completion date: 04/21  
- onset-day:=4+subclinical period  -> Depopulation starts 4 days after the emergence of I1. 
- Initial values: I1_cow=1, S_cow=8,000, S_swine=500
- Run with different $\beta$, finding the one with similar outcome 

```{r}
initial <- c(S_c = 8000, E_c = 0, I1_c = 1, I2_c = 0, S_s = 500, E_s = 0, I1_s = 0, I2_s = 0)
#initial <- c(S_c = 3*2619, E_c = 0, I1_c = 0, I2_c =3, S_s = 3*180, E_s=0, I1_s=0, I2_s=0)
```

### equation of model

```{r}
diff <- function(time, state, params){
    with(as.list(c(state, params)), {
    ## depopulation starts after 4 days  when initial I1 ha been found
    if (time > day_depopulation_start + 4)
        {a <- 1}
    else 
        {a <- 0}
    delta_swine <- min(I2_s,16)*a
    delta_cow <- min(I2_c,16-I2_s)*a
    S_c_new <- -beta_cowtocow*(I1_c+I2_c)*S_c - beta_swinetocow*(I1_s+I2_s)*S_c
    E_c_new <- beta_cowtocow*(I1_c+I2_c)*S_c + beta_swinetocow*(I1_s+I2_s)*S_c-lambda_cow*E_c
    I1_c_new <- lambda_cow*E_c -gamma_cow*I1_c
    I2_c_new <- gamma_cow*I1_c - min(delta_cow,I2_c)*a -recover_cow*I2_c
    S_s_new <- -beta_cowtoswine*(I1_c+I2_c)*S_s-beta_swinetoswine*(I1_s+I2_s)*S_s
    E_s_new <- beta_cowtoswine*(I1_c+I2_c)*S_s+beta_swinetoswine*(I1_s+I2_s)*S_s-lambda_swine*E_s
    I1_s_new <- lambda_swine*E_s -gamma_swine*I1_s
    I2_s_new <- gamma_swine*I1_s - min(delta_swine,I2_s)*a -recover_swine*I2_s
    return(list(c(S_c = S_c_new, E_c = E_c_new, I1_c = I1_c_new, I2_c = I2_c_new, 
                  S_s = S_s_new, E_s = E_s_new, I1_s = I1_s_new, I2_s = I2_s_new)))})
    }
```

## Analysis 1: Estimation of $\beta$
- How does the difference of $\beta$ contribute to final size?

from: http://lin.alic.go.jp/alic/month/domefore/2011/aug/spe-02.htm  
宮崎県における口蹄疫からの復興への取り組み  
"4月20日の発生から130日を要した口蹄疫の防疫措置は完了し、"

- Thus max_day=130 may be appropriate.
- max_day=150 from Itao-san's advice

```{r}
final_size_est <- function(params.=params_tmp, 
                           initial.=initial, 
                           max_day=150, 
                           func.=diff) {
    days <- 0:max_day
    result <- ode(y=initial., times=days, func=func., parms=params.)
    R_c <- initial["S_c"] - result[, "S_c"] - result[, "E_c"] - result[, "I1_c"] - result[, "I2_c"]
    R_s <- initial["S_s"] - result[, "S_s"] - result[, "E_s"] - result[, "I1_s"] - result[, "I2_s"]
    result <- as.data.frame(cbind(result, R_c, R_s))
    
    final_size <- result$R_c[max_day] + result$R_s[max_day] 
    return(final_size)
}
```

### power of $\beta$ 
$\beta * 10^{-3}$ - $\beta * 10^3$   

```{r}
pows <- seq(-3, 3, 0.02)
final_size_list <- c()
for (p in pows){
    params_tmp <- params_func(beta*10^p)
    final_size_list <- append(final_size_list, final_size_est())
}
    
df_final_size <- cbind.data.frame(pows, final_size_list)

plt <-  ggplot(aes(x=pows, y=final_size_list), data = df_final_size)
plt <- plt + geom_line(size = 1.5)
plt <-  plt + theme_bw()
plt <- plt + theme(axis.text=element_text(size=18), 
               axis.title=element_text(size=18,face="bold"), 
               text = element_text(size=18), 
               legend.text=element_text(size=18))
plt <- plt + xlab("Power of beta: log10") + ylab("Final size")
show(plt)

ggsave(filename="./figs/fig1-1_final_size_by_power_beta.png", plot=plt, 
       dpi=100, width=10.0, height = 8.0)
```

- $\beta * 10^p$  
- $\beta*10^{-0.20}$: `r beta * 10^(-0.20)` -> `r dplyr::filter(df_final_size, pows == "-0.2")$final_size_list`
- $\beta*10^{-0.18}$: `r beta * 10^(-0.18)` -> `r dplyr::filter(df_final_size, pows == "-0.18")$final_size_list`

###  Change of final size when $\beta$ is proportionaly changed


```{r}
beta_low <- beta * 10^(-0.20)
beta_high <- beta * 10^(-0.18)
betas <- seq(beta_low, beta_high, length.out=1000)
final_size_list <- c()

for (b in betas){
    params_tmp <- params_func(b)
    final_size_list <- append(final_size_list, final_size_est())
}
    
df_final_size <- cbind.data.frame(betas, final_size_list)

plt <-  ggplot(aes(x=betas, y=final_size_list), data = df_final_size)
plt <- plt + geom_line(size = 1.5)
## final size of actual outbreak: needs confirmation
plt <-  plt + geom_segment(x=0, xend=40, y=2000, yend=16, size=1.0)  
plt <-  plt + theme_bw()
plt <- plt + theme(axis.text=element_text(size=18), 
               axis.title=element_text(size=18,face="bold"), 
               text = element_text(size=18), 
               legend.text=element_text(size=18))
plt <- plt + xlab("beta") + ylab("Final size")
show(plt)

ggsave(filename="./figs/fig1-2_linear_beta_vs_finalsize.png", plot=plt, 
       dpi=100, width=10.0, height = 8.0)
```

- Result

```{r}
print(df_final_size[310:320, ])

## from output
beta <- 1.854162e-05
```

## Analysis 2: Final size simulation
- Initiation day of depopulationを連続的に変化させて、final sizeの振る舞いを調べた。

from: http://lin.alic.go.jp/alic/month/domefore/2011/aug/spe-02.htm  
宮崎県における口蹄疫からの復興への取り組み  
"4月20日の発生から130日を要した口蹄疫の防疫措置は完了し、"

- max_day=300 was used to better show the presence of deadline

### Depopulation from 0-60 days

```{r}
params_tmp <- params_func(beta.=beta)
depop_date <- 0:60
final_size_list <- c()
for (d in depop_date){
    params_tmp$day_depopulation_start <- d
    final_size_list <- append(final_size_list, final_size_est(max_day=300))
}
    
df_final_size <- cbind.data.frame(depop_date, final_size_list)

plt <-  ggplot(aes(x=depop_date, y=final_size_list), data = df_final_size)
plt <- plt + geom_line(size = 1.5)
plt <-  plt + theme_bw()
plt <- plt + theme(axis.text=element_text(size=18), 
               axis.title=element_text(size=18,face="bold"), 
               text = element_text(size=18), 
               legend.text=element_text(size=18))
plt <- plt + xlab("Initiation day of depopulation") + ylab("Final size")
plt 
ggsave(filename="./figs/fig2_different_initiation_date.png", plot=plt, 
       dpi=100, width=10.0, height = 8.0)
```

## Analysis 3: Sensitivity analysis of beta
- Final size simulation was done under different beta values.  
- $0.6-1.4 *\beta$ was used for simulation.  
- max_day=300 was used to better show the presence of deadline


```{r}
## This process may take some time
mults <- seq(0.6, 1.4, 0.05)
depop_date <- 0:100
df_depop_mult <- data.frame()

for (m in mults){
    params_tmp <- params_func(m*beta)
    final_size_list <- c()
        for (d in depop_date){
            params_tmp$day_depopulation_start <- d
            #max_day=130 yields strange results; need to check
            final_size_list <- append(final_size_list, final_size_est(max_day=300))
        }
    mult <- rep(m, length(final_size_list))
    df_depop <- cbind.data.frame(depop_date, final_size_list, mult)
    df_depop_mult <- rbind(df_depop_mult, df_depop)
    }

```

```{r}
plt <-  ggplot(aes(x=depop_date, y=final_size_list, group=factor(mult)), data = df_depop_mult)
plt <- plt + geom_line(aes(colour = mult), size = 1.5)
plt <- plt + scale_colour_gradient2(low="blue", mid="gray", high="red", midpoint=1.0)
plt <-  plt + theme_bw()
plt <- plt + theme(axis.text=element_text(size=18), 
               axis.title=element_text(size=18,face="bold"), 
               text = element_text(size=18), 
               legend.text=element_text(size=18))
plt <- plt + xlab("Initiation day of depopulation") + ylab("Final size")
plt 

ggsave(filename="./figs/fig3_final_size_by_beta_and_initiation_date.png", plot=plt, 
       dpi=100, width=10.0, height = 8.0)
```

## Analysis 4: Equational interpretation of this model 

- max_day=300 was used to better show the presence of deadline

```{r}
max_day <- 300
days <- 0:max_day
params = params_func(beta.=beta)

result <- ode(y=initial, times=days, func=diff, parms=params)

R_c <- initial["S_c"] - result[, "S_c"] - result[, "E_c"] - result[, "I1_c"] - result[, "I2_c"]
R_s <- initial["S_s"] - result[, "S_s"] - result[, "E_s"] - result[, "I1_s"] - result[, "I2_s"]
result <- as.data.frame(cbind(result, R_c, R_s))

result_ed <- result %>% 
    dplyr::mutate(FOI = params$beta_cowtocow * (I1_c+I2_c)*S_c + 
                      params$beta_swinetocow * (I1_s+I2_s)*S_c + 
                      params$beta_cowtoswine * (I1_c+I2_c)*S_s + 
                      params$beta_swinetoswine * (I1_s+I2_s)*S_s)
plt <- ggplot(aes(time, FOI), data = result_ed)
plt <- plt + geom_line(size = 1.5)
plt <-  plt + geom_segment(x=0, xend=80, y=16, yend=16, size=1.0)
plt <- plt + scale_x_continuous(limits = c(0, 80))
plt <-  plt + theme_bw()
plt <- plt + theme(axis.text=element_text(size=18), 
                   axis.title=element_text(size=18,face="bold"), 
                   text = element_text(size=18), 
                   legend.text=element_text(size=18))
plt <- plt + xlab("Days") + ylab("Daily New Cases")
plt

ggsave(filename="./figs/fig4_equation_interpretation.png", plot=plt, 
       dpi=100, width=10.0, height = 8.0)
```

Analysis 4: Sensitivity analysis of 牧場間の感染のパラメータ
- Final size simulation was done under different 牧場間の感染のパラメータ values.  
- How to implement the change parameters?
- No need to implement.
